---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

** MESSAGE TO THE REVIEWER **

the current Rstudefault is to save image in the **(file_name)_files** folder.
Please find the figures into **"PA1_template_files\figure-html"**

the rmarkdown::render() has been used to produce the .md and .html. knit2html is
depreciated.

```{r}
library(readr)
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data
```{r}
activity <- read_csv('activity.zip')
str(activity)
```


## What is mean total number of steps taken per day?

```{r}
sum_step_day <-
  activity %>% 
  group_by(date) %>% 
  summarise(step_day = sum(steps,na.rm = TRUE))

mean_step_day <- mean(sum_step_day$step_day, na.rm = TRUE)
median_step_day <- median(sum_step_day$step_day, na.rm = TRUE)
```

With NA removed, the median is `r as.character(median_step_day)` and the mean is `r mean_step_day` steps per day

```{r hist_with_na}

hist(sum_step_day$step_day, 
     main = 'total number of steps taken per day',
     xlab = 'number of steps')
abline(v=mean_step_day,col='blue')
text(mean_step_day - 2000, 20, 'mean', col = 'blue')
abline(v=median_step_day,col='red')
text(median_step_day + 2000, 20, 'median', col = 'red')

```

## What is the average daily activity pattern?

```{r average_day}

int_mean <-
  activity %>% 
  group_by(interval) %>% 
  summarise(mean_int_steps = mean(steps, na.rm = TRUE))

max_avg_step = 
  int_mean %>% 
  arrange(desc(mean_int_steps)) %>% 
  top_n(n=1,wt=mean_int_steps)

plot(int_mean, type='l', main='a standard day of steps'); text(x=max_avg_step$interval, y=0, as.character(max_avg_step$interval), col='red');
abline(v=max_avg_step$interval, col = 'grey')
points(max_avg_step, col = 'red')

```

Maximum number of steps in the **`r max_avg_step$interval`th** interval.


## Imputing missing values

```{r}
na_count = is.na(activity$steps) %>% sum
```

There are `r na_count` NAs in the *activity* dataset.

We are going to impute the NAs with the average value on the associated interval and save it in the **steps_imputed** variable

```{r}
activity_na_imputed <-
  activity %>%
  left_join(int_mean, by='interval') %>% 
  mutate(steps = coalesce(steps, mean_int_steps)) %>% 
  select(date, interval, steps)

```

```{r}

sum_step_day_imputed <-
  activity_na_imputed %>% 
  group_by(date) %>% 
  summarise(step_day = sum(steps,na.rm = TRUE))

mean_step_day_imputed <- mean(sum_step_day_imputed$step_day, na.rm = TRUE)
median_step_day_imputed <- median(sum_step_day_imputed$step_day, na.rm = TRUE)

diff_mean <- mean_step_day_imputed - mean_step_day
diff_median <- median_step_day_imputed - median_step_day
```

With NA imputed, the median is `r as.character(median_step_day_imputed)` and the mean is `r as.character(mean_step_day_imputed)` steps per day.
Compared with the NA remove statistics, the median is changed by `r diff_median` and the mean by `r diff_mean`.
It clearly have an impact, mostly on the mean.


```{r hist_na_imputed}
hist(sum_step_day_imputed$step_day, 
     main = 'total number of steps taken per day - imputation performed',
     xlab = 'number of steps')
abline(v=mean_step_day_imputed,col='blue',lwd=4)
text(mean_step_day_imputed - 2000, 20, 'mean', col='blue')
abline(v=median_step_day_imputed,col='red', lty=10, lwd=4)
text(median_step_day_imputed + 2000, 20, 'median', col = 'red')

```


## Are there differences in activity patterns between weekdays and weekends?


```{r workday_vs_weekend}

weekend <-
activity %>% 
  mutate(day_type = as.factor(ifelse(weekdays(date) %in% c('Samstag', 'Sonntag'), 'weekend', 'weekday'))) %>% 
  group_by(interval, day_type) %>% 
  summarise(avg_daily = mean(steps,na.rm = TRUE))

library(lattice)

xyplot(avg_daily ~ interval | day_type, data=weekend, type='l',layout = c(1,2),
       xlab='interval', ylab='number of steps', main= 'comparing the weekends and weekdays')


```

